---
title: "Predicting Wine Quality with Regression Model"
subtitle: "Prediction with Regression Model"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "Havva Nur Elveren"
output:
  html_document:
      theme: journal
      toc: yes
      toc_depth: 4
      #toc_float: true
  word_document:
      toc: yes
      toc_depth: 4
      #toc_float: true
  pdf_document:
      toc: yes
      theme: journal
      toc_depth: 4
      #toc_float: true
---
---
# Objective: Predicting Wine Quality
Can we predict wine quality based on its features such as acidity, alcohol, sugar or sulfate level? In this project, we'll predict Wine Quality with looking at the value of different features of a wine. We'll use a data set that has been collected from red wine variants of the Portuguese "Vinho Verde" wine. If quality is greater than 6.5 it is considered as good wine, otherwise it is considered as bad wine.

# Data Description:
* 1.6K Row with 12 Column. You can download the data from the link https://www.kaggle.com/uciml/red-wine-quality-cortez-et-al-2009
```{r}
library(kableExtra)

dt <- data.frame(Name = c("fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide",
"density", "pH", "sulphates", "alcohol", "quality"),
Description = c("most acids involved with wine or fixed or nonvolatile (do not evaporate readily)", "the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste", "found in small quantities, citric acid can add 'freshness' and flavor to wines", "the amount of sugar remaining after fermentation stops, it's rare to find wines with less than 1 gram/liter and wines with greater than", "the amount of salt in the wine", "the free form of SO2 exists in equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents", "amount of free and bound forms of S02; in low concentrations, SO2 is mostly undetectable in wine, but at free SO2", "the density of water is close to that of water depending on the percent alcohol and sugar content", "describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the", "a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and", "the percent alcohol content of the wine","Score between 0 and 10, if quality > 6.5 it's Good, otherwise it is Bad "))

dt %>%
  kbl() %>%
  kable_styling()
```
## Step 1: Load the Libraries
```{r}
library(tidyverse)
library(broom)
theme_set(theme_classic())

library(ISLR)
library(caret)
library(readxl)
library(pROC)
library(lattice)
library(ggplot2)
library(dplyr)
library(e1071) 
library(corrplot)
library(kknn)
library(ggplot2)
library(MLeval)
library(AppliedPredictiveModeling)
library(corrplot)
library(Hmisc)
library(dplyr)
library(ggfortify) 
library(psych)
library(MASS)
```
## Step 2: Load the Data Set and Visualize the Relationships Among Features
```{r}
winedata <- read.csv("winequality-red.csv")
winedata <- data.frame(winedata, stringsAsFactors = FALSE)
str(winedata)

table(winedata$quality)

#Visualize the correlation between features.
pairs.panels(winedata[c("fixed.acidity","volatile.acidity","citric.acid","residual.sugar","chlorides","sulphates", "alcohol","free.sulfur.dioxide","total.sulfur.dioxide","density","pH","quality")])

```

## Step 3: Prepare the Test and Training Data
```{r}
## Create random training and test data set with splitting the data with the proportion of %75 to %25.
trainIndex <- createDataPartition(winedata$quality, p = 0.75, list=FALSE)
trainData <- winedata[trainIndex,]
testData <- winedata[-trainIndex,]

prop.table(table(trainData$quality))
prop.table((table(testData$quality)))

```

## Step 4: Create Linear Regression Model
```{r}
#Create Linear Regression Model for dependent value quality and independent value sulphates
regModel1 <- lm(quality~sulphates, data=trainData)
regModel1
summary(regModel1)

# Creating Diagnostics Plots:
regModel1.diagnostics <- augment(regModel1)
head(regModel1.diagnostics)

ggplot(regModel1.diagnostics, aes(sulphates, quality))+geom_point()+ggtitle('Fig.1. Wine Quality vs Sulphate Level')+stat_smooth(method = lm, se = FALSE)+geom_segment(aes(xend = sulphates, yend = .fitted), color = "red", size = 0.3)

#Diagnostics Plots
par(mfrow = c(2, 2))
plot(regModel1)

plot(regModel1, 4)

# Predict Quality with Sulphate Level
linear_model_pred1 <- predict(regModel1, testData)
my_data=as.data.frame(cbind(predicted=linear_model_pred1,observed=testData$quality))

ggplot(my_data,aes(linear_model_pred1,testData$quality))+geom_point()+geom_smooth(method=lm)+ggtitle('Linear Regresion Model Accuracy With Sulphate Level')


#Create Linear Regression Model for dependent value quality and independent value alcohol
regModel <- lm(quality~alcohol, data=trainData)
regModel
summary(regModel)

# Creating Diagnostics Plots:
regModel.diagnostics <- augment(regModel)
head(regModel.diagnostics)

ggplot(regModel.diagnostics, aes(alcohol, quality)) +
geom_point() +
ggtitle('Fig.1. Wine Quality vs Alcohol Level')+
stat_smooth(method = lm, se = FALSE) +
geom_segment(aes(xend = alcohol, yend = .fitted), color = "red", size = 0.3)

#Diagnostics Plots
par(mfrow = c(2, 2))
plot(regModel)

plot(regModel, 4)

# Make Prediction and Plot the accuracy
linear_model_pred <- predict(regModel, testData)
my_data=as.data.frame(cbind(predicted=linear_model_pred,observed=testData$quality))

ggplot(my_data,aes(linear_model_pred,testData$quality))+geom_point()+geom_smooth(method=lm)+ggtitle('Linear Regression Model Accuracy With Alcohol Level')

```

## Step 5: Create Multiple Linear Regression Model
```{r}

control <- trainControl(method = "repeatedcv", number = 10, repeats = 5, verbose = FALSE, search = "grid")
metric <- "RMSE"
tuneLength <- 10

# Train multiple linear regression model with using all the features
Multi_Linear_Reg <- train(quality ~ ., data=trainData, method="lm", metric=metric, preProc=c("center", "scale"), trControl=control, tuneLength = tuneLength)
summary(Multi_Linear_Reg)

# Make Prediction and Plot the accuracy
Multi_linear_model_pred <- predict(Multi_Linear_Reg, testData)

rmse<-RMSE(Multi_linear_model_pred, testData$quality)
rmse

# Error Rate
error.rate=rmse/mean(testData$quality)
error.rate

my_data=as.data.frame(cbind(predicted=Multi_linear_model_pred,observed=testData$quality))

ggplot(my_data,aes(Multi_linear_model_pred,testData$quality))+geom_point()+geom_smooth(method=lm)+ggtitle('Multiple Linear Regression Accuracy')

# Print, plot variable importance
print(varImp(Multi_Linear_Reg, scale = FALSE))

plot(varImp(Multi_Linear_Reg, scale = FALSE), main="Variable Importance using LM")

```

## Step 6: Improve Model Performance
```{r}
#Add non-linear term for Alcohol, Sulphates, Volatile.Acidity, Chlorides, Total.Sulfur.Dioxide, free.sulfur.dioxide and pH

regModelWithNonLinear <- lm(quality ~ poly(alcohol,sulphates, volatile.acidity,chlorides,free.sulfur.dioxide,total.sulfur.dioxide,pH, degree = 2), data=trainData)
summary(regModelWithNonLinear)

# Select independent variables using AIC
fullModel <- lm( quality ~., data = trainData) 

stepModel <- stepAIC( fullModel, direction = "both", trace = FALSE) 

summary(stepModel)

```

## Step 7: Compare Performance Improvement Models
```{r}
anova(stepModel,regModelWithNonLinear)
```
# Conclusion
In this project we predicted wine quality with several different regression models. First we created single linear regression model to predict quality with using alcohol level and sulfates level as independent variables separately. When we compare the linear regression models "quality vs sulphates" and "quality vs alcohol", we see that r-squared value is significantly higher when use alcohol value alone, which means using alcohol as independent value gives us a more complext model. If we look at prediction accuracy plots, it also shows us quality vs alcohol gives us a better accuracy. 

If we look at the multiple linear regression model, that we include all the features into the model, With this full model, r-squared value increases to 0.3766 which is an improvement comparing to using a simple linear regression model. If we look at the variable importance graph, alcohol, sulphates, volatile.acidity,residual.sugar,chlorides, free.sulfur.dioxide, total.sulfur.dioxide and pH has significant importance when predicting wine quality.

Having this information we try to improve model performance with only using the important values and adding a non-linear term for all the important values when creating the linear regression model. And we see that r-squared value increased to  0.4297 which is significant improvement.

However selecting independent variables with AIC didn't show any improvement on the model performance.

Comparing the different regression models we applied to our data set, we can conclude that simple linear regression is not a good method to predict wine quality for this data set. Using only important variables as predictor and adding non-linear terms helps to improve the model performance. However the r-squared value that our best model produced is 0.4297, and it still not a strong prediction model for this case. 




